---
title: "Homework 2"
author: "JingYao Geng"
output: github_document
---

```{r setup, message = FALSE, echo = FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

#### **Read the Mr.Trashwheel dataset.**

```{r Alternative, eval = FALSE, echo = FALSE}
# First, define a path to the dataset.
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"

trashwheel_df = 
  read_xlsx(
    path = path_to_data,
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )

```

```{r trashwheel_df, collapse = TRUE, message = FALSE}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
trashwheel_df
```


#### **Read precipitation data for 2018 and 2017.**

**2018:**
```{r precip_2018, echo = FALSE}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% #drop the row based on the missing values in var. 'month'
  mutate(year = 2018) %>%
  relocate(year)

precip_2018
```

**2017:**

```{r precip_2017, echo = FALSE}
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>% #drop the row based on the missing values in var. 'month'
  mutate(year = 2017) %>%
  relocate(year)

precip_2017
```


**Now combine annual precipitation:**

```{r combine precip, echo = TRUE, collapse = TRUE}
# create a data frame for month
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )

# row-combine precip_2018 and precip_2017
precip_df = 
  bind_rows(precip_2018, precip_2017) 
  # Type: bind_rows(precp_2018,precp_2017) %>% view in the console to view it.

# left-join: 
precip_m = left_join(precip_df, month_df, by = "month") 
precip_m 
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the  inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation for year 2017 and 2018. The total monthly precipitation for year 2017 is `r sum(pull(precip_2017, total))` (in). The total monthly precipitation for year 2018 is `r sum(pull(precip_2018, total))` (in). 
